---
title: "Effect of quality-adjusted production function inputs"
author: "Paul Brockway, João dos Santos, Matthew Kuperus Heun"
date: "10 April 2015"
output: pdf_document
---

```{r Setup, message=FALSE, warning=FALSE}
require(EconData)
require(EconModels)
require(dplyr)
historicalData <- filter(rbind(IST, Leeds, REXS1960), Country %in% c("UK", "PT"))
dataPT <- filter(historicalData, Country == "PT")
dataPTIST <- filter(dataPT, Source == "IST")
```

This is a study of the effects of quality-adjusted production function inputs
on the importance of energy in the production function.

# Historical Data

We use the following time series for the production functions:

## Unadjusted variables

Variable | PT       | UK    
-------- | -------- | -----------
Output   | iGDP     | iGDP
Capital  | iKstkS.L | iKstkO.WwithRD
Labor    | iL       | iL
Energy   | iXpMP    | iXp

## Quality-adjusted variables

Variable | PT         | UK    
-------- | ---------- | ---------------
Output   | iGDP       | iGDP
Capital  | iKservS.L  | iKservO.WwithRD
Labor    | ihLest     | ihLest
Energy   | iUMP       | iU

# Portugal (PT) results

## Cobb-Douglas, unadjusted inputs

```{r PT-unadj}
naturalCoef(cdModel(formula = iGDP ~ iKstkS.L + iL + iYear, 
                    data = dataPTIST, constrained = TRUE))

naturalCoef(cdModel(formula = iGDP ~ iKstkS.L + iL + iXpMP + iYear, 
                    data = dataPTIST, constrained = TRUE))
```

## Cobb-Douglas, quality-adjusted inputs

```{r PT-qadj}
naturalCoef(cdModel(formula = iGDP ~ iKservS.L + ihLest + iYear, 
                    data = dataPTIST, constrained = TRUE))

naturalCoef(cdModel(formula = iGDP ~ iKservS.L + ihLest + iUMP + iYear, 
                    data = dataPTIST, constrained = TRUE))
```
