---
output: github_document
bibliography: vignettes/References.bib
---

<!-- README.md is generated from README.Rmd. Please edit that file -->

```{r, echo = FALSE}
knitr::opts_chunk$set(
  collapse = TRUE,
  comment = "#>",
  fig.path = "README-"
)
```

# MacroGrowth

The `MacroGrowth` package provides sophisticated tools for fitting macroeconomic 
growth models to data. 
`MacroGrowth` has several desirable features for economic modelers:

* Functions that fit several different macroeconomic growth models.
* Functions fit along 
boundaries of the economically-meaningful region, 
thereby ensuring that the best possible fit is found.
* Options to perform residual resampling, 
thereby providing a way to assess the variability of estimated parameters.
* A consistent formula-based interface to fitting functions.

The functions in this package were developed for and used first in 
Heun et al. [-@Heun:2017].


## Installation

You can install `MacroGrowth` from github with:

```{r gh-installation, eval = FALSE}
# install devtools if not already installed
# install.packages("devtools")
devtools::install_github("EconModels/MacroGrowth")
# To build vignettes locally, use
devtools::install_github("EconModels/MacroGrowth", build_vignettes = TRUE)
```

## More Information

Find more information, including vignettes and function documentation at 
<https://econmodels.github.io/MacroGrowth/>

## References


