---
title: "Using the MacroGrowth package"
author: "Matthew Kuperus Heun"
date: "`r Sys.Date()`"
output: rmarkdown::html_vignette
vignette: >
  %\VignetteIndexEntry{Vignette Title}
  %\VignetteEngine{knitr::rmarkdown}
  %\VignetteEncoding{UTF-8}
---

```{r setup, include = FALSE}
knitr::opts_chunk$set(
  collapse = TRUE,
  comment = "#>"
)
library(dplyr)
library(ggplot2)
library(rlang)
library(tidyr)
library(MacroGrowth)
```

## Introduction

The `MacroGrowth` package provides sophisticated tools for fitting macroeconomic 
growth models to data.
`MacroGrowth` has several desirable features for economic modelers:

* Functions that fit several different macroeconomic growth models.
* Functions that fit along 
boundaries of the economically-meaningful region, 
thereby ensuring that the best possible fit is found.
* Options to perform residual resampling, 
thereby providing a way to assess the variability of estimated parameters.
* A consistent formula-based interface to fitting functions.

The functions in this package were developed for and used first in the paper
Heun, M.K., J. Santos, P.E. Brockway, R. Pruim, T. Domingos, and M. Sakai.
"From Theory to Econometrics to Energy Policy: 
Cautionary Tales for Policymaking Using Aggregate Production Functions".
*Energies*, pp. 1-44, 10 (203), 2017.
<http://doi.org/10.3390/en10020203>.


## Macroeconomic growth models

`MacroGrowth` supports the following macroeconomic growth models:

* Single-factor model: $y = \theta \: \mathrm{e}^{\lambda t} \: x_1^m$
* Cobb-Douglas (CD): $y = \theta \: \mathrm{e}^{\lambda t} \: 
                          x_1^{\alpha_1} \, x_2^{\alpha_2} \, x_3^{\alpha_3}$
* Constant Elasticity of Substitution (CES):
          $y = \theta \: \mathrm{e}^{\lambda t} \, \left\{\delta \left[\delta_1 x_1^{-\rho_1}
                + (1-\delta_1)x_2^{-\rho_1} \right]^{\rho/\rho_1}
                + (1-\delta) x_3^{-\rho} \right\}^{-1/\rho}$, and 
* Linear Exponential (Linex): 
          $y = \theta \: x_3 \exp{\left\{ 2 a_0 \left[ 1 - \frac{x_2 + x_3}{2 x_1} \right] + 
                      a_1 \left[ \frac{x_2}{x_3} - 1 \right] \right\}}$.

In the above equations:

* $y$ is economic output, usually measured as GDP, given in data;
* $\theta$ is a fitted scale parameter;
* $\lambda$ is the Solow residual;
* $t$ is time;
* $x_1$, $x_2$, and $x_3$ are factors of production;
* $\alpha_1$, $\alpha_2$, and $\alpha_3$ are fitted output elasticities 
  for the corresponding factors of production
  in the Cobb-Douglas growth model;
* $\delta$ and $\delta_1$ are fitted share parameters for the CES growth model;
* $\rho$ and $\rho_1$ are fitted parameters related the elasticity of substitution 
  in the CES model; and
* $a_0$ and $a_1$ are fitted coefficients in the Linex production function.

In the CES model, the elasticity of substitution between the ($x_1$, $x_2$) pair and $x_3$
is given by $\sigma = \frac{1}{1 + \rho}$.
The elasticity of substitution between $x_1$ and $x_2$ is given by 
$\sigma_1 = \frac{1}{1 + \rho_1}$.

The Linex model is usually presented with $a_0$ (as above) and $c_t$ coefficients.
$c_t$ is the ratio of Linex fitted coefficients,
given by 
$c_t = \frac{a_1}{a_0}$.


## Data

The starting point for any fitting process is macroeconomic data.
For this vignette, we'll use a subset of the UK macroeconomic data 
from the "Cautionary Tales" paper by Heun et al. <http://doi.org/10.3390/en10020203>.
The data frame `EconUK` is included in the `MacroGrowth` package.

```{r}
library(dplyr)
library(ggplot2)
library(rlang)
library(tidyr)
library(MacroGrowth)
head(EconUK)
```

The `EconUK` data frame contains indexed
GDP, capital, labor, and two types of energy data
(primary exergy and useful exergy). 
All economic and energy data series are indexed by ratio to an initial year (1960),
such that all quantities are 1 in 1960.
Time is indexed by difference to the initial year, such that 1960 is year 0.
The prefix `i` indicates an indexed variable.
`iK` is indexed capital stock,
`iL` is indexed labor,
`iXp` is indexed primary exergy, and
`iXu` is indexed useful exergy.

The functions in `MacroGrowth` can fit indexed or non-indexed data.
When data are indexed as shown in the `EconUK` data frame, 
the fitted value of the $\theta$ parameter is expected to be close to 1.
When data are *not* indexed, the $\theta$ parameter will likely be far from 1,
because $\theta$ assumes the task of unit conversion 
from factors of production in various functional forms
(on the right side of the economic model)
to economic output
(on the left side of the economic model).


## Fitting examples

To fit a model to macroeconomic data, use one of the `*Model` functions
in the `MacroGrowth` pacakge.


### Single-factor model

To fit a single-factor model, use the `sfModel` function.

```{r}
sffit <- sfModel(formula = iGDP ~ iK + iYear, data = EconUK)
sffit
```

The `sfModel` function produces an `SFmodel` object.

In all `*Model` functions, 
a formula of the form `y ~ x1 + x2 + x3 + t` 
communicates the roles of the factors of production to the fitting function.
The components of the formula are: 

* the "response" variable (usually indexed GDP), positioned to the left of the `~` character,
* indexed factors of production, separated by `+` and positioned to the right of the `~` character,
and  
* the time variable, indexed to 0 at the top of the data frame,
separated from the factors of production by a `+` character, and 
always positioned as the rightmost item in the formula. 

For models with fewer than 3 factors of production, 
some of `x1`, `x2`, and `x3` can be omitted.  

In the above example,
`formula = iGDP ~ iK + iYear` indicates that
the model is to be fitted with 
indexed GDP (iGDP) as the response variable,
indexed capital stock (iK) as the only factor of production, and
indexed year (iYear) as the time variable.

The coefficients of the fitted model can be recovered with the `naturalCoef` function.

```{r}
naturalCoef(sffit)
```


### Cobb-Douglas model

Similar functions can be used for the other model types. 
To fit the Cobb-Douglas model to the UK's capital stock and labor data,
use the `cdModel` function.

```{r}
cdfit <- cdModel(formula = iGDP ~ iK + iL + iYear, data = EconUK)
cdfit
```

The `cdModel` function produces a `CDEmodel` object.

In the above example, `iGDP ~ iK + iL + iYear` indicates that
the model is to be fitted with 
indexed GDP (iGDP) as the response variable,
indexed capital stock (iK) and indexed labor (iL) as the factors of production, and
indexed year (iYear) as the time variable.

The coefficients of the model can be recovered with the `naturalCoef` function.

```{r}
naturalCoef(cdfit)
```

For the CES production function, use the `cesModel` function 
with two or three factors of production.

```{r}
cesfit <- cesModel(formula = iGDP ~ iK + iL + iYear, data = EconUK)
naturalCoef(cesfit)
```

The `cesModel` function produces a `cesModel` object.

Note that output elasticities ($\alpha_1$, $\alpha_2$, and $\alpha_3$) 
are calculated by the `naturalCoef` function.
Equations for the output elasticities are given in Appendix B 
of the "Cautionary Tales" paper by Heun et al. <http://doi.org/10.3390/en10020203>.


### Linex model

For the Linex model, 
there must be three factors of production.
The first factor of production ($x_1$) must be capital stock.
The second factor of production ($x_2$) must be labor.
The third factor of production ($x_3$) must be an energy variable.

```{r}
linexfit <- linexModel(formula = iGDP ~ iK + iL + iXp + iYear, data = EconUK)
naturalCoef(linexfit)
```

The `linexModel` function produces a `LINEXmodel` object.

To recover the output elasticities from the Linex model, 
use the `fortify` function.

```{r}
######## fortify is not found ? ######
# fortify(linexfit)
```

### Metadata

By default the original economic data are saved 
as an attribute of the model object.

```{r}
head(cdfit$data)
```
To prevent this behavior, set the `save.data` argument to `FALSE`.


### Fitted values

The fitted values of the response variable ($y$) can be recovered with the `yhat` function.
The `yhat` function gives values of the response variable 
on the natural (not logarithmic) scale.

```{r}
yhat(cdfit)
```

A graph comparing fitted and historical values can be created as shown below.

```{r, fig.width = 6, fig.align = "center"}
EconUK %>%
  bind_cols(yhat(cdfit) %>% as.data.frame() %>% set_names("yhat")) %>% 
  ggplot() + 
  geom_point(mapping = aes(x = Year, y = iGDP), shape = 1) +
  geom_line(mapping = aes(x = Year, y = yhat))

```


## Accounting for boundaries and constraints

The `*Model` fitting functions have the capability to account for boundaries and constraints.


### Single-factor model

The simplest fitting function (`sfModel`)
provides an option to constrain the fit with $m = 1$,
thereby assuming a Cobb-Douglas-type model where the output elasticity 
is constrained to the economically-meaningful value.
By default, the single-factor model sets `constrained = FALSE`.

```{r}
sfModel(formula = iGDP ~ iK + iYear, data = EconUK) %>% naturalCoef()
sfModel(formula = iGDP ~ iK + iYear, data = EconUK, constrained = TRUE) %>% naturalCoef()
```


### Cobb-Douglas model

The `cdModel` function always assumes constant returns to scale and
fits with $\displaystyle\sum_i \alpha_i = 1$.
By default, `constrained = TRUE`, thereby assuring that $0 \le \alpha_i \le 1$.

To demonstrate fitting with constraints, 
we first use the `cdModel` function to fit the first 10 years (1960--1969) of `EconUK`
and find $\alpha_2 < 0$.

```{r}
cdModel(formula = iGDP ~ iK + iL + iXu + iYear, data = EconUK %>% filter(Year < 1970), constrained = FALSE) %>% naturalCoef()
```

If, on the other hand, we use the default setting for the `constrained` argument,
the best fit is found with $\alpha_2 = 0$ and different values for `\alpha_1` and `\alpha_3`.

```{r}
cdModel(formula = iGDP ~ iK + iL + iXu + iYear, data = EconUK %>% filter(Year < 1970)) %>% naturalCoef()
```


### CES model

Constraints on the CES model are considerably more complicated than constraints on single-factor or Cobb-Douglas models.
When `constrained = TRUE`, the `cesModel` function restricts 
fitted parameters to the following economically-meaningful regions:

* $0 \le \delta, \, \delta_1 \le 1$
* $-1 \le \rho, \, \rho_1 < \infty$ (and, therefore, $0 \le \sigma, \, \sigma_1 < \infty$)

The default value of `constrained` for the `cesModel` function is `TRUE`.

To demonstrate, we fit the CES model unconstrained 
with capital (`iK`) and useful exergy (`iXu`) as the factors of production
for the years 1980--1989
to find that $\rho_1$ is beyond the economically meaningful boundary.

```{r}
cesModel(formula = iGDP ~ iK + iXu + iYear, data = EconUK %>% filter(Year >= 1980 & Year < 1990), constrained = FALSE) %>% naturalCoef()
```

If, instead, we use the default value for `constrained` (`TRUE`), 
we obtain fitted parameters that fall within the economically-meaningful region
or on the boundary of the economically-meaningful region,
even if the quality of the fit is worse.

```{r}
cesModel(formula = iGDP ~ iK + iXu + iYear, data = EconUK %>% filter(Year >= 1980 & Year < 1990)) %>% naturalCoef()
```

The process for constrained fitting by the `cesModel` function is described in detail 
in the "Cautionary Tales" paper by Heun et al. <http://doi.org/10.3390/en10020203>.
In short, the full CES model and 20 boundary models are compared. 
The model with the smallest sum of squared errors (SSE) is returned by the `cesModel` function.
Models with larger SSE are stored as a list in the `model.attempts` attribute of the `cesModel` object
returned by the `cesModel` function.
To retrieve the rejected boundary models, use the `attr` function.

```{r}
attr(cesfit, "model.attempts")[[1]] %>% naturalCoef()
```


### Linex model

The economically-meaningful region for Linex model parameters $a_0$ and $a_1$ is all real numbers.
Thus, fitted parameter constraints have no meaning for the `linexModel` function, and 
`constrained` is not an argument to the `linexModel` function.


## Resampling

## Plotting results

## References

* cesEst/Henningsen and Henningsen
* Cautionary tales paper



